Designing and Developing a Personalised Recommender System

### 2021-02-11 paper/paper.tex
rewriting evaluation methods

\subsection{Data description}
The data for my recommender system was generated from the Yelp Open Dataset \cite{YelpDataset}. 
This is a subset of all the businesses, reviews, and user data on Yelp, formatted as JSON files. 
The dataset includes approximately 8,000,000 reviews of 200,000 businesses by 2,000,000 users, as well as 
additional Covid-19 related features of each business such as whether they now offer takeout or delivery. 
There are 3 sections of the dataset, with 5 JSON files in the main dataset, 1 in the photos section, and 
1 in the covid section. 
The $business.json$ file contains business data including location data, attributes, and categories. 
$review.json$ contains full review text data including the $user\_id$ that wrote the review and the $business\_id$ 
that the review is written for. 
The file $user.json$ holds every user's first name along with other data adn metadata which I don't use \cite{YelpDataset}. 
The data files $checkin.json$ and $tip.json$ are not used in my system. 
The dataset also includes approximately 200,000 photos but these are not relevant to my application, since 
the interface is command line only, so they were discarded. 
The covid related data from $covid\_features.json$ contains additional information about each business with regards to 
the measures they have taken due to the pandemic, such as offering delivery or takeout options or a 
special message for customers. 
Every business in the dataset is in 1 of 10 metropolitan areas in the USA and Canada 
(Montreal, Calgary, Toronto, Pittsburgh, Charlotte, Urbana-Champaign, Phoenix, Las Vegas, Madison, and Cleveland) \cite{YelpDataset}.

\subsection{Data preparation and feature selection}
The entire Yelp dataset contains a lot of data which is not relevant to my domain of restaurants, so much data 
preparation is required to save computation time and in order to provide more accurate predictions. 
The dataset is provided in JSON format, but in such a way that every entry is its own JSON on its own line, 
which does not lend well to fast lookups by the main program. 
Therefore it was necessary to convert every file I used in the system into a typical JSON file, such that it could 
be loaded as a python dictionary with the lookup key being either the uniquely identifying $business\_id$ or 
$user\_id$ depending on the case. 
This allows for extremely fast data lookups in the system. 
It was of course also necessary to filter the list of all Yelp businesses down to just the restaurants. 
I further narrowed the list of restaurants by selecting only those which fit into one or more of the top 12 
categories in terms of number of restaurants (American, Mexican, Italian, Chinese, Seafood, Japanese, Canadian, 
Mediterranean, Indian, Thai, Middle Eastern, Vietnamese, in that order). 
Since the user selects one of these 12 options for categories of restaurants they want to find, they will 
never be recommended a restaurant which doesn't fit at least one of these categories. 
I chose to use the top 12 categories because it's enough to provide a wide variety of options to the user without 
being so many that the user is overwhelmed by the amount. 
I also ensured that every restaurant was still open and had at least 1 check-in in 2019 (the last year that 
data was available). 
This left 20,565 restaurants which could be recommended to the user, which needed to be split into their respective 
metropolitan areas by using the latitude and longitude values of the businesses and of the centres of each city. 
All other datasets were filtered to only data pertaining to these restaurants, with duplicates in the 
covid dataset removed. 
The reviews were further filtered to only those which had 5 or more 'useful' votes by other users. 
This was to ensure I was only using high quality reviews which others found helpful, as well as to allow the 
recommender system to compute results in a reasonable amount of time (~10 seconds on my machine). 
This left around 120,000 reviews, which were then converted to a CSV file ($reviews.csv$) with unnecessary data 
removed and structured as follows: ($User\_ID$, $Business\_ID$, $Rating$).

@@

\subsection{Evaluation methods}
How to evaluate

@@

\subsection{Recommendation algorithm}
The collaborative-filtering aspect of the algorithm is carried out by first converting the ratings data to a 
matrix with $User\_ID$ as the rows and $Business\_ID$ as the columns. 
If a user has not rated a restaurant then the rating in that cell is set to 0. 
The prediction $\widehat{r}_{ui}$ (user $u$'s rating of item $i$) is set as:
\begin{equation}
    \widehat{r}_{ui} = \mu + b_u + b_i + q_i^Tp_u
\end{equation}
Where the parameters are learnt by minimising the following regularised squared error: 
\begin{equation}
    \sum_{r_{ui}\in{R_{train}}} (r_{ui} - \widehat{r}_{ui})^2 + \lambda(b_i^2 + b_u^2 + ||q_i||^2 + ||p_u||^2)
\end{equation}
This minimisation is performed by stochastic gradient descent:
\begin{align}
    \begin{split}
    & b_u \gets b_u + \gamma(e_{ui} - \lambda b_u) \\
    & b_i \gets b_i + \gamma(e_{ui} - \lambda b_i) \\ 
    & p_u \gets p_u + \gamma(e_{ui} \cdot q_i - \lambda p_u) \\ 
    & q_i \gets q_i + \gamma(e_{ui} \cdot p_u - \lambda q_i) \\
    \end{split}
\end{align}
where $e_{ui} = r_{ui} - \widehat{r}_{ui}$. 
These steps are performed over all the ratings of the trainset and repeated 20 times \cite{SVD}.

The idea is to find restaurants in a new city which are similar to those in another city that the user likes, just like Entree. 
I used a model-based approach to collaborative-filtering based on matrix factorisation. 
The specific model is Single Value Decomposition (SVD), which was popularised by Simon Funk during the Netflix Prize. 
In my program, I fit the algorithm on a training set, and then for every restaurant of the specified category 
and in the specified city, I predict what the user would rate that restaurant. 
Then I sort those restaurants by their score and take the top 8 to present to the user.

@@

\subsection{Comparison against baseline implementation}
Compare vs generic suggestions

\subsection{Comparison against hybrid recommenders in related studies}
Read some papers

@@

\subsection{Further developments}
What could I do in the future \cite{5284958}. 
Implement the idea of particular categories being closer some than others. 
Specify price range and/or consider price when comparing similarities of restaurants. 
Include demographic information

>>>>

\subsection{Data description}
The data for my recommender system was generated from the Yelp Open Dataset \cite{YelpDataset}. 
This is a subset of all the businesses, reviews, and user data on Yelp, formatted as JSON files. 
The dataset includes approximately 8,000,000 reviews of 200,000 businesses by 2,000,000 users, as well as 
additional Covid-19 related features of each business such as whether they now offer takeout or delivery. 
There are 3 sections of the dataset, with 5 JSON files in the main dataset, 1 in the photos section, and 
1 in the covid section. 
The $business.json$ file contains business data including location data, attributes, and cuisine. 
$review.json$ contains full review text data including the $user\_id$ that wrote the review and the $business\_id$ 
that the review is written for. 
The file $user.json$ holds every user's first name along with other data adn metadata which I don't use \cite{YelpDataset}. 
The data files $checkin.json$ and $tip.json$ are not used in my system. 
The dataset also includes approximately 200,000 photos but these are not relevant to my application, since 
the interface is command line only, so they were discarded. 
The covid related data from $covid\_features.json$ contains additional information about each business with regards to 
the measures they have taken due to the pandemic, such as offering delivery or takeout options or a 
special message for customers. 
Every business in the dataset is in 1 of 10 metropolitan areas in the USA and Canada 
(Montreal, Calgary, Toronto, Pittsburgh, Charlotte, Urbana-Champaign, Phoenix, Las Vegas, Madison, and Cleveland) \cite{YelpDataset}.

\subsection{Data preparation and feature selection}
The entire Yelp dataset contains a lot of data which is not relevant to my domain of restaurants, so much data 
preparation is required to save computation time and in order to provide more accurate predictions. 
The dataset is provided in JSON format, but in such a way that every entry is its own JSON on its own line, 
which does not lend well to fast lookups by the main program. 
Therefore it was necessary to convert every file I used in the system into a typical JSON file, such that it could 
be loaded as a python dictionary with the lookup key being either the uniquely identifying $business\_id$ or 
$user\_id$ depending on the case. 
This allows for extremely fast data lookups in the system. 
It was of course also necessary to filter the list of all Yelp businesses down to just the restaurants. 
I further narrowed the list of restaurants by selecting only those which fit into one or more of the top 12 
cuisines in terms of number of restaurants (American, Mexican, Italian, Chinese, Seafood, Japanese, Canadian, 
Mediterranean, Indian, Thai, Middle Eastern, Vietnamese, in that order). 
Since the user selects one of these 12 options for cuisines they want to find, they will 
never be recommended a restaurant which doesn't fit at least one of these categories. 
I chose to use the top 12 cuisines because it's enough to provide a wide variety of options to the user without 
being so many that the user is overwhelmed by the amount. 
I also ensured that every restaurant was still open and had at least 1 check-in in 2019 (the last year that 
data was available). 
This left 20,565 restaurants which could be recommended to the user, which needed to be split into their respective 
metropolitan areas by using the latitude and longitude values of the businesses and of the centres of each city. 
All other datasets were filtered to only data pertaining to these restaurants, with duplicates in the 
covid dataset removed. 
The reviews were further filtered to only those which had 5 or more 'useful' votes by other users. 
This was to ensure I was only using high quality reviews which others found helpful, as well as to allow the 
recommender system to compute results in a reasonable amount of time (~10 seconds on my machine). 
This left around 120,000 reviews, which were then converted to a CSV file ($reviews.csv$) with unnecessary data 
removed and structured as follows: ($User\_ID$, $Business\_ID$, $Rating$).

@@

\subsection{Evaluation methods}
The evaluation of the system was done using offline experiments. 
The accuracy of ratings predictions was evaluated using Mean Absolute Error (MAE), which is calculated by the 
following equation:
\begin{equation}
    MAE = \frac{1}{|\mathcal{T}|} \sum_{(u,i)\in{\mathcal{T}}} {|\widehat{r}_{ui} - r_{ui}|}
\end{equation}
This metric is appropriate for this RS since it can handle large errors on a few items, which may occur in this system 
due to some rare skewed ratings. 
The MAE of this system is around 0.99. 
Measuring the accuracy of usage predictions is done using precision, calculated by: 
\begin{equation}
    \frac{TP}{TP+FP}
\end{equation}
where TP and FP correspond to True Positive and False Positive respectively. 
This is because the number of recommendations is predefined and not very large (8 items). 
The precision of this system is 0.125, since 1 out of the 8 items is a true positive. 
Novelty, diversity, coverage

@@

\subsection{Recommendation algorithm}
The collaborative-filtering aspect of the algorithm is carried out by first converting the ratings data to a 
matrix with $User\_ID$ as the rows and $Business\_ID$ as the columns. 
If a user has not rated a restaurant then the rating in that cell is set to 0. 
The prediction $\widehat{r}_{ui}$ (user $u$'s rating of item $i$) is set as:
\begin{equation}
    \widehat{r}_{ui} = \mu + b_u + b_i + q_i^Tp_u
\end{equation}
Where the parameters are learnt by minimising the following regularised squared error: 
\begin{equation}
    \sum_{r_{ui}\in{R_{train}}} (r_{ui} - \widehat{r}_{ui})^2 + \lambda(b_i^2 + b_u^2 + ||q_i||^2 + ||p_u||^2)
\end{equation}
This minimisation is performed by stochastic gradient descent:
\begin{align}
    \begin{split}
    & b_u \gets b_u + \gamma(e_{ui} - \lambda b_u) \\
    & b_i \gets b_i + \gamma(e_{ui} - \lambda b_i) \\ 
    & p_u \gets p_u + \gamma(e_{ui} \cdot q_i - \lambda p_u) \\ 
    & q_i \gets q_i + \gamma(e_{ui} \cdot p_u - \lambda q_i) \\
    \end{split}
\end{align}
where $e_{ui} = r_{ui} - \widehat{r}_{ui}$. 
These steps are performed over all the ratings of the trainset and repeated 20 times \cite{SVD}. 
The knowledge-based aspect of the algorithm produces recommendations based on the explicit location and cuisine 
preferences of the user. 
These are then cascaded down to the CF algorithm, which assigns each a predicted rating based on the ratings data. 
The 8 recommendations with the highest predicted rating are then presented to the user.

@@

\subsection{Comparison against baseline implementation}
Cross validation with 5 folds was carried out on both the SVD system and a baseline implementation. 
The 5 Mean Absolute Errors for the SVD system had a mean of 0.9962 and a standard deviation of 0.0033. 
The MAE results for the baseline implementation had a mean of 1.0091 and a standard deviation of 0.0056. 
This shows that when compared to a baseline implementation, the SVD system does perform better, generating more 
accurate predictions, however the difference is not particularly large.

\subsection{Comparison against hybrid recommenders in related studies}
The Entree restaurant recommender system is an example of a cascade style hybrid recommender system which 
uses both knowledge-based and collaborative-filtering methods to help users find a restaurant they might like in 
a city they've never been to before \cite{burke1999integrating,burke2000knowledge,burke2002hybrid,burke2007hybrid}. 
This is very similar to the way my system works, with the biggest difference being the vastly greater amount of 
knowledge engineering that has gone into Entree. 
This additional knowledge is used in Entree, and other FindMe style systems, to give more accurate similarity assessments 
of restaurants by considering factors such as price and quality of experience as well as location and cuisine, 
and by arranging these factors hierarchically. 
For example, Entree arranges these as: cuisine, price, quality of experience, others, but other systems allow the 
user to specify this order themselves. 
This is part of the reason why Entree would outperform this system.

@@

\subsection{Further developments}
What could I do in the future \cite{5284958}. 
Implement the idea of particular cuisines being closer some than others. 
Specify price range and/or consider price when comparing similarities of restaurants. 
Include demographic information
